---
title: "p8105_hw5_yz4990"
author: "Yucheng Zhao"
date: "2024-11-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
#library(rvest)



```

## Problem 1
function to check for same birthdays
```{r}
birthday_check = function(group_size) {
  birthdays <- sample(1:365, size = group_size, replace = TRUE)
  duplicate = duplicated(birthdays)
  
  for (x in duplicate) {
    if (x == TRUE) {
      return(TRUE)
    }
  }
  return(FALSE)
}
```

1000 simulations and plot
```{r}
sim_1000 = 
  expand_grid(
    iter = 1:1000, 
    group_size = 2:50
  ) |> 
  mutate(
    same_birthday = map_lgl(group_size, \(x) birthday_check(x))
  ) |> 
  group_by(group_size) |> 
  summarize(prob = mean(same_birthday))
  
  sim_1000 |> 
  ggplot(aes(x = group_size, y = prob)) +
  geom_point() + 
  labs(
    title = "Group Size vs. Probability of Having Same Birthday",
    x = "Group Size",
    y = "Probability"
  ) +
  theme_minimal()
```


## Problem 2
function for simulation and one-sample t-test
```{r}
n = 30
sd = 5
mu = 0

sim_t_test = function(n = 30, mu, sd = 5, num_sim = 5000) {
  sim_data = expand_grid(
    iter = 1:num_sim,
    sample_size = n
  ) |> 
    mutate(
      x_df = map(sample_size, \(x) tibble(rnorm(x, mu, sd))), 
      t_test = map(x_df, \(x) broom::tidy(t.test(x)))
    ) |> 
    unnest(t_test)
  
  return(sim_data)
}

sim_5000 = sim_t_test(n = 30, mu = 0, sd = 5, num_sim = 5000) |> 
  select(estimate, p.value)

sim_5000

```


power, average estimated mean, average estimated mean of the rejected
```{r}
mu = c(1, 2, 3, 4, 5, 6)

sim_results = map_dfr(mu, \(x) sim_t_test(n = n, mu = x, sd = sd, num_sim = 5000) |> 
  mutate(true_mean = x)) |> 
  group_by(true_mean) |> 
  summarize(
    power = mean(p.value < 0.05), 
    avg_est_mu = mean(estimate), 
    avg_mu_rejected = mean(estimate[p.value < 0.05]),
    .groups = "drop"
  )
```

Power vs True Mean
```{r}
sim_results |> 
  ggplot(aes(x = true_mean, y = power)) +
  geom_point() +
  labs(
    title = "Power vs True Mean",
    x = "True Mean",
    y = "Power"
  ) +
  theme_minimal()
```

As the effect size increases, the power also increases. When the true mean of the sample is close to 0, the power is less because the difference between the sample mean and the null hypothesis is small and hard to detect by the t-test. As the true mean becomes larger, the power approaches to 1 because the difference between the sample mean and the null hypothesis is large and easy to detect.


True Mean vs Average Estimated Mean
```{r}
sim_results |> 
  ggplot(aes(x = true_mean, y = avg_est_mu)) +
  geom_point() +
  labs(
    title = "True Mean vs Average Estimated Mean",
    x = "Average Estimated Mean",
    y = "True Mean"
  ) +
  theme_minimal()

```


```{r}
sim_results |> 
  ggplot(aes(x = true_mean, y = avg_mu_rejected)) +
  geom_point() +
  labs(
    title = "True Mean vs Average Estimated Mean of the Rejected",
    x = "True Mean",
    y = "Average Estimated Mean of the Rejected"
  ) +
  theme_minimal()


```

The average estimated mean of the rejected is closer or approximately equal to the true mean when the true mean is large, and it deviates from the true mean when the true mean is small. This is because there are more cases where the null hypothesis mu = 0 is rejected as the true mean increases.


## Problem 3
```{r}
homi_df = read_csv("./data/homicide-data.csv", 
                   na = c("NA", "Unknown", ""), 
                   ) |> 
  janitor::clean_names()

```

The raw data has `r ncol(homi_df)` columns X `r nrow(homi_df)` rows.  The variables are `r colnames(homi_df)`. There are some missing values in the raw data, noted as "Unknown" or an empty space. 

proportion test for Baltimore
```{r}
homi_df = homi_df |> 
  mutate(
    city_state = paste(city, state, sep = ", ")
    ) |> 
  group_by(city_state) |> 
  summarize(
    total_homi = n(),
    unsolved = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )

baltimore_df = homi_df |> 
  filter(city_state == "Baltimore, MD")

baltimore_test = prop.test(
  pull(baltimore_df, unsolved), 
  pull(baltimore_df, total_homi)
) |> 
  broom::tidy() |> 
  select(estimate, conf.low, conf.high)

baltimore_test
```

proportion test for all cities and plots
```{r}
cities_test = homi_df |> 
  mutate(
    prop_test = map2(unsolved, total_homi, \(x, y) prop.test(x, y) |> 
                       broom::tidy())
  ) |> 
  unnest(prop_test) |> 
  select(city_state, estimate, conf.low, conf.high)

  cities_test |> 
    mutate(city_state = forcats::fct_reorder(city_state, estimate)) |> 
    ggplot(aes(x = city_state, y = estimate)) +
    geom_point() +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
    theme_minimal() +
  labs(
    title = "Estimated Proportion of Unsolved Homicides in Cities with 95% CI",
    x = "City",
    y = "Proportion"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

```









```{r}


```









